Variance reduction for Markov chains via martingale representations

### 2020-04-13 aap/Revision/vr-mcmc-revision-new-new.tex
update vr-mcmc-revision-new-new.tex

\pi_{n}^{N}(f)=\frac{1}{n}\sum_{p=N+1}^{N+n} f(X^x_{p}),

@@

&=& \frac{1}{n}\sum_{k=1}^{K}\sum_{l=N+1}^{N+n}\sum_{p=0}^{N+n-l}\bar{a}_{p,k}(X_{l-1}^{x})\phi_{k}(\xi_{l}).

@@

Then we solve   the least squares optimization problems

@@

\widetilde{A}_{q,k} \in \argmin_{\psi \in \Psi} \sum_{s = 1}^{T} \left| \sum_{r = 0}^{q}  f(X_{r+l_0}^{(s)})\phi_k(\xi^{(s)}_{l_0+1}) - \psi(X^{(s)}_{l_0})\right|^2

@@

for $k=1,\ldots,K,$ and some $l_0$ with $l_0+q\leq N+n,$
where \(\Psi\) is a  class of  real-valued functions on \(\mathbb{R}^d\) and $\xi^{(s)}_{1},\ldots,\xi^{(s)}_{N+n}$ are iid random variables with distribution $P_\xi$ used to construct the $s$-th trajectory of $X.$   As usual in regression analysis (see Section~11 in \cite{gyorfi2006distribution}),  we  truncate  the least squares estimates $(\widetilde{A}_{q,k})$ and set

@@

\widehat M_{K,n}^N(f) := \frac{1}{n}\sum_{k=1}^{K}\sum_{l=N+1}^{N+n}\left[\widehat{A}_{N+n-l,k}(X_{l-1}^{x})\phi_{k}(\xi_{l})\right].

@@

In the next section we analyze the case of ULA and show that under some rather mild conditions (convexity of the potential $U$ outside a ball around zero)

@@

\PE[A^2_{l,k}(X^x_{N+l-1})]\leq e^{\varkappa n\gamma^2}R^2_k ,\quad k=1,2,\ldots, \quad l=0,\ldots,n-1,

@@

\bar A_{q,k}:=\argmin_{\psi\in\Psi}\mathsf{E}[|A_{q,k}(X_{l_0})-\psi(X_{l_0})|]^2

@@

\PVar[\widehat \pi_{K,n}^N(f)|\TrainSet]\leq \frac{2Ce^{\varkappa n\gamma^2}}{n}.

@@

\frac{\mathrm{cost}(\widehat \pi_{K,n}^N) \PVar[\widehat \pi_{K,n}^N(f)|\TrainSet]}{\mathrm{cost}(\pi_{n}^N)\PVar[\pi_{n}^N(f)]}\leq C e^{\varkappa n\gamma^2} K \gamma \, T^2 \mathtt{D}_\Psi.

@@

A^2_{q,\mathbf{k}}(x)\leq C_Ke^{\varkappa n\gamma^2}(1+V(x)),\quad 1\leq |\mathbf{k}|\leq K,

@@

\sum_{\|\mathbf{k}\|\geq K+1} A^2_{q,\mathbf{k}}(x)\leq C_{K}e^{\varkappa n\gamma^2}\sum_{I\subseteq\{1,\ldots,d\},\, I\neq \emptyset}

>>>>

\pi_{n}^{N}(f)= \frac{1}{n} \sum_{p=N+1}^{N+n} f(X^x_{p}),

@@

&=& \frac{1}{n}\sum_{k=1}^{K}\sum_{l=N+1}^{N+n}
A_{N+n-1,k}(X_{l-1}^{x}) \phi_{k}(\xi_{l}),

@@

Then we solve  for $k=1,\ldots,K,$  the least squares optimization problems

@@

\widetilde{A}_{q,k} \in \argmin_{\psi \in \Psi} \sum_{s = 1}^{T} \left| \sum_{r = 0}^{q}  f(X_{r+N}^{(s)})\phi_k(\xi^{(s)}_{N+1}) - \psi(X^{(s)}_{N})\right|^2

@@

where \(\Psi\) is a  class of  real-valued functions on \(\mathbb{R}^d\) and $\xi^{(s)}_{1},\ldots,\xi^{(s)}_{N+n}$ are i.i.d. random variables with distribution $P_\xi$ used to construct the $s$-th trajectory of $X.$   As usual in regression analysis (see \cite[Section~11]{gyorfi2006distribution}),  we  truncate  the least squares estimates $(\widetilde{A}_{q,k})$ and set

@@

\widehat M_{K,n}^N(f) := \frac{1}{n}\sum_{k=1}^{K}\sum_{l=N+1}^{N+n} \widehat{A}_{N+n-l,k}(X_{l-1}^{x})\phi_{k}(\xi_{l})\,.

@@

In the next section we analyze the case of ULA algorithm and show that under some rather mild conditions (convexity of the potential $U$ outside a ball around zero)

@@

\PE[A^2_{l,k}(X^x_{N+l-1})]\leq \rme^{\varkappa n\gamma^2}R^2_k ,\quad k \in \nset, \quad l=0,\ldots,n-1,

@@

\bar A_{q,k}:=\argmin_{\psi\in\Psi}\mathsf{E}[|A_{q,k}(X_{N})-\psi(X_{N})|^2]

@@

\PVar[\widehat \pi_{K,n}^N(f)|\TrainSet]\leq \frac{2C \rme^{\varkappa n\gamma^2}}{n}.

@@

\frac{\mathrm{cost}(\widehat \pi_{K,n}^N) \PVar[\widehat \pi_{K,n}^N(f)|\TrainSet]}{\mathrm{cost}(\pi_{n}^N)\PVar[\pi_{n}^N(f)]}\leq C \rme^{\varkappa n\gamma^2} K \gamma \, T^2 \mathtt{D}_\Psi.

@@

A^2_{q,\mathbf{k}}(x)\leq C_K \rme^{\varkappa n\gamma^2}(1+V(x)),\quad 1\leq |\mathbf{k}|\leq K,

@@

\sum_{\|\mathbf{k}\|\geq K+1} A^2_{q,\mathbf{k}}(x)\leq C_{K} \rme^{\varkappa n\gamma^2}\sum_{I\subseteq\{1,\ldots,d\},\, I\neq \emptyset}
